---
title: The Golem of Praque
subtitle: Practice
author: Max Hachemeister
date: last-modified
toc: true
mainfont: DejaVu Sans
fontsize: 12pt
format:
  gfm:
    toc: true
  typst:
    toc: true
    toc-depth: 4
number-sections: false

---

```{r}
#| label: setup
#| output: false

library(tidyverse)
library(ggplot2)
library(rethinking)
```


### 2E4.
> What does it mean to say: "The probability of water is '0.7'."?

Probability does not exist. So a water probability of 0.7 does not
mean that the globe's actual proportion of water is 0.7. It means more
something like: "Given what I know, I am 70% certain that I will
see water with the next globe toss."; or "In 70% of the globe tosses
water will be observed."

### 2M1.
> Compute and plot the grid approximate posterior distribution for each
> of the following sets of observations. In each case, assume a uniform
> prior for $p$.
> 
>  1. W, W, W
>  2. W, W, W, L
>  3. L, W, W, L, W, W, W
  
This is to be done with programming.

#### This is the workflow as copied from the book.

First, calculate the values:
```{r}
#define grid
p_grid <- 
  seq(from = 0, to = 1, length.out = 20)

#define prior
prior <- 
  rep(1, 20)

#compute likelihood at each value in grid
likelihood <- 
  dbinom(6, size = 9, prob = p_grid)

#compute compute product of likelihood and prior
unstd.posterior  <- 
  likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- 
  unstd.posterior / sum(unstd.posterior)

```

Then plot:
```{r}
# Put it in tibble
data <- 
  tibble(
    p_grid,
    prior,
    likelihood,
    unstd.posterior,
    posterior
  )

data |> 
  ggplot(aes(p_grid, posterior)) +
  geom_point() +
  geom_line()
```
#### This is the solution to the task

Now, what needs to be changed in the above cases are the values for the binomial distribution according to the sample observed. As we have defined water as *success* and we have observed "W, W, W",  we tell the function that there were three successes out of a samplesize of three, and then ask how likely this observation would have been, given the actual distribution of water and land were each to the values in the `p_grid`. So we ask, what's the most likely actual distribution of water and land, that the observed sample could have resulted from.

Let's do them sequentually and maye you will see.
I will reuse the tibble we already made, and just change the `likelihood`

##### W, W, W
```{r}
data |> 
  mutate(
    likelihood = dbinom(3, size = 3, prob = p_grid)
  ) |> 
  ggplot(aes(p_grid, posterior)) +
  geom_point() +
  geom_line()
```
Ah well, the joke's on me. I need to have the other values recomputed also. So while I think I could write that up in a function, I will do it copy & paste but do the tibble and plot in the same step

```{r}
# make it a tibble
tibble(
  p_grid           = seq(from = 0, to = 1, length.out = 20),
  prior            = rep(1, 20),
  # this here is the thing that changes
  likelihood       = dbinom(3, size = 4, prob = p_grid),
  unstd.posterior  =  likelihood * prior,
  posterior        = unstd.posterior / sum(unstd.posterior)
) |> 
  #plot it
  ggplot(aes(p_grid, posterior)) +
  geom_point() +
  geom_line()
```

##### W, W, W, L

```{r}
# make it a tibble
tibble(
  p_grid           = seq(from = 0, to = 1, length.out = 20),
  prior            = rep(1, 20),
  # this here is the thing that changes
  likelihood       = dbinom(3, size = 4, prob = p_grid),
  unstd.posterior  =  likelihood * prior,
  posterior        = unstd.posterior / sum(unstd.posterior)
) |> 
  #plot it
  ggplot(aes(p_grid, posterior)) +
  geom_point() +
  geom_line()
```


##### L, W, W, L, W, W, W
```{r}
# make it a tibble
tibble(
  p_grid           = seq(from = 0, to = 1, length.out = 20),
  prior            = rep(1, 20),
  # this here is the thing that changes
  likelihood       = dbinom(5, size = 2, prob = p_grid),
  unstd.posterior  =  likelihood * prior,
  posterior        = unstd.posterior / sum(unstd.posterior)
) |> 
  #plot it
  ggplot(aes(p_grid, posterior)) +
  geom_point() +
  geom_line()
```

### 2M2.